Windowed residual-minimization model reduction for dynamical systems

### 2019-10-03 numerical_experiments.tex

\section{Numerical Experiments}\label{sec:numerical_experiments}
This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity. In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs. Additionally, we assess the impact of the window size, time-step size, and time scheme on \methodAcronym. It is noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs. Future state prediction and/or interpolation to new flow conditions is not considered as these problems introduce confounding factors (e.g., accuracy of the basis).

\subsection{Sod Shock Tube}
We first consider reduced-order models of the Sod shock tube problem. The system is governed by the compressible Euler equations in one dimension, 
\begin{equation}\label{eq:euler_1D}
    \frac{\partial \boldsymbol u}{\partial t} + \frac{\partial \boldsymbol F}{\partial \mathsf{x}} = 0, \quad
    \boldsymbol u= 
    \begin{Bmatrix} \rho \\ \rho u \\ \rho E \end{Bmatrix}, \quad 
    \boldsymbol F = \begin{Bmatrix} \rho u \\ \rho u^2 + p \\  u(\rho E + p) \end{Bmatrix},
\end{equation}
where $\boldsymbol u : \Omega \times [0,T] \rightarrow \RR{3}$ comprise the density, x-momentum, and total energy, $\mathsf{x} \in \Omega \defeq  [0,1]$ is the spatial domain, and 
$T = 1$ the final time. 
The problem setup is given by the initial conditions,

@@

\subsubsection{Description of FOM and Generation of Trial Space}
The 1D compressible Euler equations are solved using a finite volume method. The domain is partitioned into 500 cells of uniform width. The finite volume method uses the Rusanov flux~\cite{rusanov} at the cell interfaces. The FOM uses the Crank--Nicolson (CN) scheme, which is a linear multistep method defined by the coefficients $\alpha_0 = 1,\alpha_1 = -1, \beta_0 = \beta_1 = 1/2$, for temporal integration. The FOM is evolved for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$. The \spatialAcronym\ trial subspace is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $. The resulting trial subspace comprises $99.99\%$ of the snapshot matrix energy.
%  The solution is saved every other time-step, resulting in 500 solutions snapshots. These snapshots are used to generate a trial basis for the reduced-order models through POD. The trial basis is selected to be of dimension $K = 46$, which corresponds to a $99.99\%$ energy criterion. No affine transformation is used in the construction of the trial subspace, i.e., $\stateIntercept = \bz$.

\subsubsection{Description of Reduced-Order Models}
Reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approaches are considered. No hyper-reduction is considered 
in this example, i.e., $\stweightingMatArg{} = \mathbf{I}$. Details on the implementation of the 
different reduced-order models is as follows:
\begin{itemize}
\item \textbf{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
a constant time-step size of $ \Delta t = 0.002$.

@@

\subsubsection{Numerical Results}
The impact of the window size on \methodAcronymROMs\ is first assessed. A set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$ are considered. The standard Galerkin and LSPG ROMs are additionally considered. Results are shown for the direct method with the CN time-discretization scheme; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section. 
First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = \{0.5,1.0\}$. Figure~\ref{fig:sod_xt} shows 
$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, it is observed that the LSPG and Galerkin ROMs accurately characterize 
the system: they correctly track the shock location, expansion waves, etc. Both predictions, however, are observed to by highly oscillatory. These oscillations are 
not physical and can lead to numerical blowup; e.g., due to negative pressure. The \methodAcronymROMs\ produce less oscillatory solutions than both 
the Galerkin and LSPG ROMs. Critically, it is seen that, as the window size over which the residual is minimized grows, the solution becomes less oscillatory. 
When the residual is minimized over the entire space--time domain, the solution displays no oscillations.

@@

\subsubsection{Summary of Numerical Results for the Sod Shock Tube}
The key observations from the results of the first numerical example are: 
\begin{enumerate}
\item Increasing the window size over which the residual is minimized lead to more physically relevant solutions. Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ lead to less oscillatory solutions.
\item Increasing the window size over which the residual is minimized does not necessary lead to a more accurate solution in the $\elltwo$ sense. We observed that minimizing the residual over an intermediary window size led to the most accurate results as measured by the $\elltwo$ norm.
\item The \methodAcronym\ displays time-step convergence: both the integrated $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time-step size decreased. This is in contrast to LSPG.
\item For all examples considered, in where the windows comprised up to 2000 time-steps, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
direct method was observed to be, in general, slightly more efficient and robust than the indirect method.
\end{enumerate}

>>>>

\section{Numerical experiments}\label{sec:numerical_experiments}
This section analyzes the performance of \methodAcronymROMs\ leveraging \spatialAcronym\ trial subspaces on two benchmark problems: the Sod shock tube and compressible flow in a cavity. In each experiment, \methodAcronym\ is compared to the Galerkin and LSPG ROMs. Additionally, we assess the impact of the window size, time-step size, and time scheme on \methodAcronym. It is noted that both experiments are designed to test the \textit{reproductive} ability of the ROMs. Future state prediction and/or interpolation to new flow conditions is not considered as these problems introduce confounding factors (e.g., accuracy of the basis).

\subsection{Sod shock tube}
We first consider reduced-order models of the Sod shock tube problem. The system is governed by the compressible Euler equations in one dimension, 
\begin{equation}\label{eq:euler_1D}
    \frac{\partial \boldsymbol u}{\partial t} + \frac{\partial \boldsymbol F}{\partial \mathsf{x}} = 0, \quad
    \boldsymbol u= 
    \begin{Bmatrix} \rho \\ \rho u \\ \rho E \end{Bmatrix}, \quad 
    \boldsymbol F = \begin{Bmatrix} \rho u \\ \rho u^2 + p \\  u(\rho E + p) \end{Bmatrix},
\end{equation}
where $\boldsymbol u : \Omega \times [0,T] \rightarrow \RR{3}$ comprise the density, x-momentum, and total energy, $\mathsf{x} \in \Omega \defeq  [0,1]$ is the spatial domain, and 
$T = 1$ the final time. 
The problem setup is given by the initial conditions,

@@

\subsubsection{Description of FOM and generation of trial space}
The 1D compressible Euler equations are solved using a finite volume method. The domain is partitioned into 500 cells of uniform width. The finite volume method uses the Rusanov flux~\cite{rusanov} at the cell interfaces. The FOM uses the Crank--Nicolson (CN) scheme, which is a linear multistep method defined by the coefficients $\alpha_0 = 1,\alpha_1 = -1, \beta_0 = \beta_1 = 1/2$, for temporal integration. The FOM is evolved for $t \in [0.0,1.0]$ at a time-step of $\Delta t = 0.002$. The \spatialAcronym\ trial subspace is constructed by executing Algorithm~\ref{alg:pod} with inputs $N_{\text{skip}}=2, \stateIntercept = \bz, K=46 $. The resulting trial subspace comprises $99.99\%$ of the snapshot matrix energy.
%  The solution is saved every other time-step, resulting in 500 solutions snapshots. These snapshots are used to generate a trial basis for the reduced-order models through POD. The trial basis is selected to be of dimension $K = 46$, which corresponds to a $99.99\%$ energy criterion. No affine transformation is used in the construction of the trial subspace, i.e., $\stateIntercept = \bz$.

\subsubsection{Description of reduced-order models}
Reduced-order models based on the Galerkin, LSPG, and \methodAcronym\ approaches are considered. No hyper-reduction is considered 
in this example, i.e., $\stweightingMatArg{} = \mathbf{I}$. Details on the implementation of the 
different reduced-order models is as follows:
\begin{itemize}
\item \textbf{Galerkin ROM}: The Galerkin ROM is obtained through Galerkin projection of the FOM. The Galerkin ROM is evolved in time with the CN time scheme at 
a constant time-step size of $ \Delta t = 0.002$.

@@

\subsubsection{Numerical results}
The impact of the window size on \methodAcronymROMs\ is first assessed. A set of \methodAcronymROMs\ that minimize the residual over windows of constant size 
$\DeltaSlabArg{n} \equiv \DeltaSlabArg{} =  .002$, $0.004$ ,$0.008$, $0.02$, $0.04$, $0.10$, $0.20$, $1.0$ are considered. The standard Galerkin and LSPG ROMs are additionally considered. Results are shown for the direct method with the CN time-discretization scheme; a comparison of different time-marching methods and direct/indirect solution techniques will be provided later in this section. 
First, Figure~\ref{fig:sod_density} presents the density solutions predicted by the various ROMs at $t = \{0.5,1.0\}$. Figure~\ref{fig:sod_xt} shows 
$x-t$ diagrams for the same density solutions. From Figures~\ref{fig:sod_density} and~\ref{fig:sod_xt}, it is observed that the LSPG and Galerkin ROMs accurately characterize 
the system: they correctly track the shock location, expansion waves, etc. Both predictions, however, are observed to by highly oscillatory. These oscillations are 
not physical and can lead to numerical blowup; e.g., due to negative pressure. The \methodAcronymROMs\ produce less oscillatory solutions than both 
the Galerkin and LSPG ROMs. Critically, it is seen that, as the window size over which the residual is minimized grows, the solution becomes less oscillatory. 
When the residual is minimized over the entire space--time domain, the solution displays no oscillations.

@@

\subsubsection{Summary of numerical results for the Sod shock tube}
The key observations from the results of the first numerical example are: 
\begin{enumerate}
\item Increasing the window size over which the residual is minimized lead to more physically relevant solutions. Specifically, we observed that as the window size over which the residual was minimized grew, \methodAcronym\ lead to less oscillatory solutions.
\item Increasing the window size over which the residual is minimized does not necessary lead to a more accurate solution in the $\elltwo$ sense. We observed that minimizing the residual over an intermediary window size led to the most accurate results as measured by the $\elltwo$ norm.
\item The \methodAcronym\ displays time-step convergence: both the integrated $\elltwo$ state error and residual norm decreased and displayed time-step convergence as the time-step size decreased. This is in contrast to LSPG.
\item For all examples considered, in where the windows comprised up to 2000 time-steps, \methodAcronym\ with the direct method is between 1x and 10x the cost of the LSPG. The 
direct method was observed to be, in general, slightly more efficient and robust than the indirect method.
\end{enumerate}

### 2019-10-03 section3/general_form.tex

\section{Windowed Least-Squares Formulation}\label{sec:tclspg} This section
outlines the proposed \methodNameLower\ 
formulation (\methodAcronym). In contrast to (1.) the Galerkin method, which minimizes the 
time-instantaneous residual, (2.) the LSPG method, which sequentially minimizes the discrete 
residual over a single time step, and (3.) the ST-LSPG method, which minimizes the discrete 
space--time residual, the proposed \methodAcronym\ method sequentially minimizes the 
time-continuous FOM ODE residual over a series of
\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
In this section, we start by outlining the \methodAcronym\ formulation  
for a general trial subspace. We then examine the \spatialAcronym\ 
trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
associated with the minimization problems.
 %The proposed method can be formulated equivalently from
%two viewpoints. The first formulation corresponds to a sequence of
%minimization problems for the generalized coordinates that minimize the FOM
%ODE residual over each time slab. We refer to this as the 
%\textit{residual-minimization viewpoint}. The second viewpoint corresponds to
%formulating an optimal control problem for the Galerkin ROM. In this
%viewpoint, the \methodAcronym\ method is formulated as an optimal control
%problem in where the objective is to find a ``controller" that minimizes the
%residual of the Galerkin ROM. We refer to this viewpoint as the
%\textit{optimal control viewpoint.}  This section outlines the derivation of the \methodAcronym\ method
%from both the residual minimization and optimal control viewpoints.

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Residual Minimization Formulation} 
We start by partitioning the
temporal domain into a set of $\nslabs$ windows of size $\DeltaSlabArg{n}$. This partitioning is depicted in Fig.~\ref{fig:slab_fig}.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} \caption{Decomposition of time
domain into time windows.} \label{fig:slab_fig} \end{centering} 
\end{figure}
Over each time window, we introduce the space--time trial subspace,
$$\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,$$ 
where $\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
\RR{\fomdim}$
is the \methodAcronym\ approximation to the state over the $n$th time window, $\timeSpaceArg{n}$ is the set of functions acting on $[\timeStartArg{n},\timeEndArg{n}]$, and  
$\timeStartArg{n}$ and $\timeEndArg{n}$ are the beginning and end of the $n$th
time window, respectively. Over each window, define the objective,
%\begin{equation}\label{eq:obj} \mathcal{J}^n(\stateArg{}{t}) = \frac{1}{2}
%\int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\state}(\tau) -
%\velocity(\stateArg{}{\tau}) \big]^T \stweightingMatArg{n}(\tau) \big[
%\dot{\state}(\tau) - \velocity(\stateArg{}{\tau}) \big] d \tau,
%\end{equation}
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RR{}, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ } \in \RR{\fomdim \times \fomdim}$ is a
symmetric positive semi-definite matrix that, e.g., enables hyper-reduction and/or modifies the norm. For simplicity, 
it is assumed that the weighting matrix is the same for all time windows.

>>>>

\section{Windowed residual minimization formulation}\label{sec:tclspg} This section
outlines the proposed \methodNameLower\ 
formulation (\methodAcronym). In contrast to (1.) the Galerkin method, which minimizes the 
time-instantaneous residual, (2.) the LSPG method, which sequentially minimizes the discrete 
residual over a single time step, and (3.) the ST-LSPG method, which minimizes the discrete 
space--time residual, the proposed \methodAcronym\ method sequentially minimizes the 
time-continuous FOM ODE residual over a series of
\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
In this section, we start by outlining the \methodAcronym\ formulation  
for a general trial subspace. We then examine the \spatialAcronym\ 
trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
associated with the minimization problems.
 %The proposed method can be formulated equivalently from
%two viewpoints. The first formulation corresponds to a sequence of
%minimization problems for the generalized coordinates that minimize the FOM
%ODE residual over each time slab. We refer to this as the 
%\textit{residual-minimization viewpoint}. The second viewpoint corresponds to
%formulating an optimal control problem for the Galerkin ROM. In this
%viewpoint, the \methodAcronym\ method is formulated as an optimal control
%problem in where the objective is to find a ``controller" that minimizes the
%residual of the Galerkin ROM. We refer to this viewpoint as the
%\textit{optimal control viewpoint.}  This section outlines the derivation of the \methodAcronym\ method
%from both the residual minimization and optimal control viewpoints.

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Residual minimization formulation} 
We start by partitioning the
temporal domain into a set of $\nslabs$ windows of size $\DeltaSlabArg{n}$. This partitioning is depicted in Fig.~\ref{fig:slab_fig}.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} \caption{Decomposition of time
domain into time windows.} \label{fig:slab_fig} \end{centering} 
\end{figure}
Over each time window, we introduce the space--time trial subspace,
$$\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,$$ 
where $\approxstate^n :  [\timeStartArg{n} , \timeEndArg{n}] \rightarrow
\RR{\fomdim}$
is the \methodAcronym\ approximation to the state over the $n$th time window, $\timeSpaceArg{n}$ is the set of functions acting on $[\timeStartArg{n},\timeEndArg{n}]$, and  
$\timeStartArg{n}$ and $\timeEndArg{n}$ are the beginning and end of the $n$th
time window, respectively. Over each window, define the objective,
%\begin{equation}\label{eq:obj} \mathcal{J}^n(\stateArg{}{t}) = \frac{1}{2}
%\int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\state}(\tau) -
%\velocity(\stateArg{}{\tau}) \big]^T \stweightingMatArg{n}(\tau) \big[
%\dot{\state}(\tau) - \velocity(\stateArg{}{\tau}) \big] d \tau,
%\end{equation}
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RR{}, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneArg{ }^T \stweightingMatOneArg{ } \in \RR{\fomdim \times \fomdim}$ is a
symmetric positive semi-definite matrix that, e.g., enables hyper-reduction and/or modifies the norm. For simplicity, 
it is assumed that the weighting matrix is the same for all time windows.

### 2019-10-03 section3/space_time.tex

\subsection{\spaceTimeAcronym\ Trial Spaces}
We now briefly consider \spaceTimeAcronym\ trial subspaces. Over 
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
with $\text{Range}(\stbasisArgnt{n}) + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

@@

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary Conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be derived using standard calculus. Differentiating the objective with respect to $\stgenstatey$ and setting equal to zero yields,
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.

>>>>

\subsection{\spaceTimeAcronym\ trial spaces}
We now briefly consider \spaceTimeAcronym\ trial subspaces. Over 
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
with $\text{Range}(\stbasisArgnt{n}) + \stateInterceptArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

@@

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be derived using standard calculus. Differentiating the objective with respect to $\stgenstatey$ and setting equal to zero yields,
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.
